\section*{Preface}
\addcontentsline{toc}{section}{Preface}%

This work is not a direct continuation of the project assignment done in the fall of 2019. The project assignemnt was done for Kongsberg Digital, and consisted of a literature review on the use of time-series clustering for automatic classification of wind turbines. Some of the theory, and some of the models of time-series clustering were transferable.

\section*{Acknowledgements}

The IMPROVE study is an ongoing cardiology study. The dataset used in this thesis is a small subset of what has been collected for the IMPROVE study. The recruitment of participants has been done from seven hospitals in Norway, and a total of 3100 patients are to be included. Thor Edvardsen, Kristina Haugaa, Harald Brunvand have organized the study, and have been in charge. Daniela Melichova og Thuy Mi Nguyen have been responsible for collecting ultrasound data from participants. Ivar Mjåland Salte is the clinician that has annotated the data. Thanks you all, without the IMPROVE study this thesis would not exist. \bigskip

My supervisor Lasse Løvstakken has helped shape the outlook of this thesis, he has answered all my questions and provided me with invaluable guidence. You have my sincerest thanks, and I hope you are able to give other students the same experience with their master thesis that I had. Andreas Østvik helped me navigate through the IMPROVE dataset, and introduced me to the many software tools developed, a big thanks to you. Thank you to Benjamin Nedregaard who provided the architecture for the neural network applied in this thesis.

\clearpage
